Principal Component Analysis of Population Genetic Data

### 2019-10-06 paper-bedpca.tex
add figures projection 1000G

\usepackage{dsfont}
\usepackage[warn]{textcomp}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{graphicx}
\graphicspath{{../figures/}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\subsection{Outlier detection in PCA}

@@

\subsection{Subsetted version of 1000G data [RENAME?]}

@@

To project PCs of a reference dataset (e.g.\ 1000G) to a target genotype data, we implement the following 3 steps in function \texttt{bed\_projectPCA} of package bigsnpr:
\begin{enumerate}
\item matching the variants of each dataset, including removing ambiguous alleles [A/T] and [C/G], and matching strand and direction of the alleles;
\item computing PCA of the reference dataset using the matched variants only;
\item projecting computed PCs to the target data using an optimised implementation of the Online Augmentation, Decomposition, and Procrustes (OADP) transformation \cite[]{zhang2019fast} described in Supplementary Materials.
\end{enumerate}

@@

@@

@@

[CAN COMPUTE SVD OF $V'[1{:}K,~]$ FROM LOW-RANK UPDATE?]

@@

\begin{figure}[!htpb]
%\centerline{\includegraphics[width=0.8\textwidth]{AUC-simu2.pdf}}
\caption{}
\label{fig:AUC-simu2}
\end{figure}

@@

% latex table generated in R 3.5.2 by xtable 1.8-4 package
% Wed Jun 19 15:19:38 2019
\begin{table}[!htpb]
\caption{AUC values on the test set for simulations with less well imputed variants (mean [95\% CI] from $10^4$ bootstrap samples). These results are plotted in figure \ref{fig:AUC-simu2}.\label{tab:AUC-simu2}}
\vspace*{0.5em}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
Scenario & stdCT & maxCT & SCT & lassosum \\ 
\hline
100  & 77.4 [76.0-78.8] & 83.9 [83.4-84.4] & 83.1 [82.6-83.6] & 80.1 [79.5-80.8] \\ 
10K  & 69.4 [68.4-70.5] & 73.0 [72.5-73.4] & 72.9 [72.5-73.3] & 71.2 [70.6-71.7] \\ 
1M   & 64.0 [63.6-64.4] & 64.0 [63.6-64.4] & 62.7 [62.3-63.0] & 64.1 [63.3-64.8] \\ 
2chr & 70.0 [68.8-71.2] & 74.4 [73.6-75.2] & 78.5 [77.9-79.1] & 73.2 [72.5-73.8] \\ 
err  & 67.0 [66.0-68.1] & 68.6 [67.7-69.5] & 69.5 [68.9-70.1] & 65.6 [64.9-66.3] \\ 
HLA  & 74.8 [72.9-76.3] & 75.3 [73.5-76.9] & 76.4 [74.5-78.0] & 75.8 [74.2-77.2] \\ 
\hline
\end{tabular}
\end{table}

>>>>

\usepackage{dsfont}
\usepackage[warn]{textcomp}
\usepackage{adjustbox}
\usepackage{multirow}
\usepackage{graphicx}
\graphicspath{{figures/}}
\DeclareMathOperator*{\argmin}{\arg\!\min}

@@

\subsection{Outlier detection in PCA [SPLIT IN MULTIPLE SECTIONS]}

@@

\subsection{Subsetted version of 1000G data [RENAME?]\label{1000G}}

@@

\subsection{Projecting PCs from a reference dataset\label{proj}}

To project PCs of a reference dataset (e.g.\ 1000G) to a target genotype data, we implement the following 3 steps in function \texttt{bed\_projectPCA} of package bigsnpr:
\begin{enumerate}
\item matching the variants of each dataset, including removing ambiguous alleles [A/T] and [C/G], and matching strand and direction of the alleles;
\item computing PCA of the reference dataset using the matched variants only;
\item projecting computed PCs to the target data using an optimised implementation of the Online Augmentation, Decomposition, and Procrustes (OADP) transformation \cite[]{zhang2019fast}, which we describe in Supplementary Materials.
\end{enumerate}

@@

We use 60\% of individuals in the 1000G data (section \ref{1000G}) to compute the PCA ($K=20$ PCs). Then, we project the remaining 40\% individuals using two methods: 1/ simply multiplying the genotypes of these individuals by the previously computed loadings, and 2/ the OADP projection (section \ref{proj}). 
When simply projecting using loadings, there is no apparent shrinkage for PC1 and PC2, a weak [REWORD?] shrinkage for PC3 and PC4, and a large shrinkage for PC5 to PC8 (Figure \ref{fig:proj1000G}).
In contrast, there is no apparent shrinkage when projecting new individuals with OADP (Figure \ref{fig:proj1000G}).
Simple projection is affected even more by this shrinkage for PC9 to PC20, while OADP appears free of this issue (Figure \ref{fig:proj1000G-2}).

\begin{figure}[!htpb]
\centerline{\includegraphics[width=0.8\textwidth]{proj1000G-PC1-8.pdf}}
\caption{Principal Component (PC) scores 1 to 8 of the 1000 Genomes project.
Black points are the 60\% individuals used for computing PCA.
Red points are the 40\% remaining individuals, projected by simply multiplying their genotypes by the corresponding PC loadings.
Blue points are the 40\% remaining individuals, projected using the Online Augmentation, Decomposition, and Procrustes (OADP) transformation.
\label{fig:proj1000G}}
\end{figure}

@@

\renewcommand{\thefigure}{S\arabic{figure}}
\setcounter{figure}{0}
\renewcommand{\thetable}{S\arabic{table}}
\setcounter{table}{0}

@@

@@

\begin{figure}[!htpb]
\centerline{\includegraphics[width=0.8\textwidth]{proj1000G-PC9-20.pdf}}
\caption{Principal Component (PC) scores 9 to 20 of the 1000 Genomes project.
Black points are the 60\% individuals used for computing PCA.
Red points are the 40\% remaining individuals, projected by simply multiplying their genotypes by the corresponding PC loadings.
Blue points are the 40\% remaining individuals, projected using the Online Augmentation, Decomposition, and Procrustes (OADP) transformation.
\label{fig:proj1000G-2}}
\end{figure}

@@

% latex table generated in R 3.5.2 by xtable 1.8-4 package
% Wed Jun 19 15:19:38 2019
\begin{table}[!htpb]
\caption{\label{tab:AUC-simu2}}
\vspace*{0.5em}
\centering
\begin{tabular}{|l|c|c|c|c|}
\hline
Scenario & stdCT & maxCT & SCT & lassosum \\ 
\hline
100  & 77.4 [76.0-78.8] & 83.9 [83.4-84.4] & 83.1 [82.6-83.6] & 80.1 [79.5-80.8] \\ 
10K  & 69.4 [68.4-70.5] & 73.0 [72.5-73.4] & 72.9 [72.5-73.3] & 71.2 [70.6-71.7] \\ 
1M   & 64.0 [63.6-64.4] & 64.0 [63.6-64.4] & 62.7 [62.3-63.0] & 64.1 [63.3-64.8] \\ 
2chr & 70.0 [68.8-71.2] & 74.4 [73.6-75.2] & 78.5 [77.9-79.1] & 73.2 [72.5-73.8] \\ 
err  & 67.0 [66.0-68.1] & 68.6 [67.7-69.5] & 69.5 [68.9-70.1] & 65.6 [64.9-66.3] \\ 
HLA  & 74.8 [72.9-76.3] & 75.3 [73.5-76.9] & 76.4 [74.5-78.0] & 75.8 [74.2-77.2] \\ 
\hline
\end{tabular}
\end{table}
